\section{Benchmark Data Design}
\label{sec:benchmark}

This section describes how the benchmark datasets were prepared: how the ideas were authored, how they were stratified by length, and how each idea was placed on a three-tier novelty ladder. The design has three goals. First, it keeps the benchmark balanced across corpora, originality labels, and input lengths, so that aggregate accuracy is less likely to be driven by class imbalance or by unusually verbose inputs. Second, it aligns each idea with the writing style of its target corpus, because scientific papers, patents, and software repositories expose different kinds of evidence. Third, it defines a labeling scheme before evaluation, which makes the results interpretable as a test of originality assessment rather than a loose relevance-matching exercise.
 
\subsection{Benchmark Dataset Overview}
 
We constructed three independent benchmark datasets, one per retrieval corpus, totaling 135 labeled cases (45 per corpus). Each dataset was designed to reflect the natural idea style of its corresponding corpus, and all cases were manually authored and labeled. Rather than sampling from existing labeled repositories or relying on automated generation, we crafted each idea to probe specific behaviors of the originality assessment pipeline -- including its ability to distinguish stronger conceptual changes from standard domain extensions.
 
Each corpus contains equal representation across the three classes (15 cases per class per corpus), reducing the risk that overall accuracy is dominated by any single class and making per-class metrics more informative. A dataset skewed toward \textit{already\_exists} -- a likely outcome of random sampling from many real corpora -- could inflate overall accuracy without providing meaningful signal about novelty detection, which is the primary use case of the system.

Table~\ref{tab:dataset} summarizes the composition of the three datasets. Each corpus is fully balanced across the three originality classes and the three length tiers introduced below (five cases per class--length cell), and spans a diverse set of application domains so that accuracy is not driven by a single topic.

\begin{table}[width=\textwidth,cols=6,pos=ht]
    \caption{Benchmark dataset composition. Each corpus contains 45 manually authored and labeled cases, balanced across three originality classes and three length tiers (5 cases per class--length cell).}
    \label{tab:dataset}
    \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lccccl@{}}
        \toprule
        Corpus & Cases & Per class & Per tier & Domains & Representative domains \\
        \midrule
        OpenAlex (papers)  & 45 & 15 & 15 & 11 & AI/ML, Biomedical, Quantum \\
        GitHub (repos)     & 45 & 15 & 15 & 19 & DevOps, Testing, Blockchain \\
        Patents (Google)   & 45 & 15 & 15 & 20 & Autonomous vehicles, Finance, Healthcare \\
        \midrule
        Total & 135 & 45 & 45 & --- & --- \\
        \bottomrule
    \end{tabular*}
\end{table}

Table~\ref{tab:datastats} reports descriptive statistics of the idea text itself. Because every stage of the pipeline consumes the raw idea string, its length is a relevant covariate: the three length tiers map onto clearly separated token bands (roughly 35, 67, and 143 tokens on average for short, medium, and long ideas), confirming that the stratification produces distinct input-size regimes rather than overlapping ones.

\begin{table}[width=\textwidth,cols=6,pos=ht]
    \caption{Descriptive statistics of the 135 benchmark ideas. Token counts are approximate (characters/4). Class balance is fixed by design at 15 \textit{already\_exists} / 15 \textit{incremental} / 15 \textit{novel} per corpus.}
    \label{tab:datastats}
    \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lccccc@{}}
        \toprule
        Corpus & Cases & Class balance & Mean tokens & Median & Range \\
        \midrule
        OpenAlex & 45  & 15/15/15 & 89 & 75 & 28--172 \\
        GitHub   & 45  & 15/15/15 & 74 & 58 & 27--147 \\
        Patents  & 45  & 15/15/15 & 74 & 64 & 26--146 \\
        \midrule
        All      & 135 & 45/45/45 & 79 & 66 & 26--172 \\
        \bottomrule
    \end{tabular*}
\end{table}

\subsection{Idea Length Stratification}
 
A key design dimension in this benchmark is \textbf{idea length}. Within each class and each corpus, ideas are evenly distributed across three length tiers: \textit{short} (one to two sentences), \textit{medium} (three to four sentences), and \textit{long} (five or more sentences), with five cases per tier per class. This stratification tests whether the pipeline's performance is sensitive to the amount of detail provided in the input description.
 
The motivation for this design is that the pipeline processes the enriched idea description at multiple stages -- query variant generation, candidate selection, and per-criterion LLM scoring all depend on the textual content of the input. A short idea provides less signal for retrieval and a narrower context for the scoring agents. A long idea gives the retrieval layer more angles to search from and gives the Layer~1 agent more material to compare against each candidate. By balancing class labels across length tiers, we can better separate the effect of description length from the effect of originality class, and estimate how much the pipeline benefits from additional input detail.
 
In real use, the follow-up question agent and idea enrichment step are intended to partially compensate for short inputs by eliciting additional context before retrieval begins. The benchmark deliberately tests raw idea text without this enrichment step, so the length effect measured here should be interpreted as an estimate for the unenriched-input setting.
 
\subsection{Corpus-Specific Idea Design}
 
A core design principle was stylistic alignment: ideas in each dataset were written to match the natural framing conventions of the corpus against which they would be retrieved. This alignment was intended to make retrieval quality and scoring reflect the target deployment setting rather than a vocabulary mismatch between the idea description and the indexed corpus. The three corpus-specific framings were defined as follows:
 
\textbf{OpenAlex (Scientific Literature).} Ideas were framed as research proposals and methodological contributions, consistent with the academic writing style of OpenAlex-indexed papers. Descriptions emphasize problem formulation, proposed mechanisms, and expected contributions -- the vocabulary a researcher would use when describing a novel method or system to peers.
 
\textbf{GitHub (Software Repositories).} Ideas were framed as practical software projects, consistent with the repository descriptions and README summaries found on GitHub. Descriptions are implementation-oriented, referencing specific technologies, frameworks, and functional behaviors rather than abstract mechanisms.
 
\textbf{Patents (Google Patents).} Ideas were framed as system-level inventions, consistent with the technical and functional language of patent abstracts. Descriptions emphasize novel system architectures, operating patterns, and distinguishing technical claims rather than experimental contributions or code-level implementations.
 
\subsection{Labeling Scheme: The Novelty Ladder}
 
All cases were labeled under a strict three-tier scheme, which we refer to as the novelty ladder. The scheme was designed to distinguish ideas that already exist in the target corpus, ideas that extend existing work, and ideas that introduce a clearly new contribution:
 
\textbf{Already Exists.} The idea is widely known, commonly implemented, or straightforwardly retrievable in the target corpus. A case receives this label if it would likely be recognized by a practitioner in the field as existing work. Representative examples include:
 
\begin{itemize}
    \item \textit{[GitHub, Testing, short]} ``A JavaScript testing setup using Jest and React Testing Library for unit tests, component tests, mocks, and coverage reports.'' (score: 14)
    \item \textit{[OpenAlex, AI/ML, short]} ``A deep residual network trained on ImageNet that uses skip connections between convolutional layers to enable stable training of very deep image classification models.'' (score: 9)
    \item \textit{[Patents, Healthcare, long]} ``An electronic health record system that stores patient data and allows secure sharing between healthcare providers. The system maintains medical history, diagnoses, medications, lab results, imaging reports, and clinical notes in a centralized or interoperable digital format. Access controls and audit logs are used to protect sensitive patient information.'' (score: 17)
\end{itemize}
 
\textbf{Incremental.} The idea shares its core concept with existing work but introduces a clear, limited improvement -- a better component, an added feature, a new application domain, or a modest architectural modification. The key criterion is that the improvement is a natural extension of prior work rather than a new mechanism. Combinations of the form ``existing method X applied to domain Y'' are classified as incremental regardless of whether the specific combination has been published. Representative examples include:
 
\begin{itemize}
    \item \textit{[GitHub, Blockchain, long]} ``A smart contract auditing tool that combines symbolic execution results with LLM-generated explanations of potential vulnerabilities. The symbolic engine identifies suspicious execution paths while the LLM converts findings into developer-friendly explanations and suggested fixes.'' (score: 41)
    \item \textit{[OpenAlex, Biomedical, long]} ``A multimodal survival prediction model for lung adenocarcinoma that combines whole-slide histopathology images with bulk RNA sequencing profiles. A cross-attention module learns relationships between visual morphology and molecular signatures.'' (score: 38)
    \item \textit{[Patents, Manufacturing, long]} ``A predictive maintenance system that learns machine-specific degradation curves by fine-tuning a cross-factory foundation model on local sensor history. The base model captures general failure patterns across many factories, while local adaptation accounts for equipment-specific wear behavior.'' (score: 56)
\end{itemize}
 
\textbf{Novel.} The idea introduces a new mechanism, control approach, or representational approach that is not easily reduced to a recombination of existing components. Novel cases are intended to change how a problem is approached, not merely what is applied to it. We excluded superficial novelty -- ideas that appear original through buzzword recombination (e.g., ``ML + X,'' ``blockchain-based Y'') but do not introduce a new underlying principle. Representative examples include:
 
\begin{itemize}
    \item \textit{[GitHub, Testing, long]} ``A software testing framework that generates tests by simulating adversarial future maintainers who misunderstand the codebase in different ways. The framework models likely human misreadings of function names, abstractions, and architectural boundaries, then creates tests that protect against bugs introduced by those misunderstandings.'' (score: 82)
    \item \textit{[OpenAlex, AI/ML, long]} ``A continual learning system where forgetting is treated as a structured compression process. After each task, the model reorganizes old knowledge into fractal attractors in weight space, leaving sparse regions available for new tasks rather than preventing forgetting through regularization or replay.'' (score: 83)
    \item \textit{[Patents, Smart Home, long]} ``A home automation system that constructs a causal model of resident behavior from passive sensor streams and uses counterfactual reasoning to infer unstated goals. It asks what outcome the resident was likely trying to achieve under different device states, then proposes new automation rules that would have produced that outcome with less effort.'' (score: 93)
\end{itemize}
 
This distinction between incremental and novel was the most consequential design decision. A stricter interpretation of novelty -- requiring a new mechanism or approach rather than a new combination -- produces a labeling scheme that is closer to expert review practice than to keyword matching.
 
\subsection{Deliberate Inclusion of Borderline Cases}
 
Each dataset includes a proportion of borderline cases -- ideas that sit close to the incremental/novel boundary and would reasonably invite disagreement among human annotators. This design choice reflects the practical difficulty of originality assessment: the boundary between a creative extension and a new mechanism is often ambiguous, and a useful assessment tool should behave meaningfully in this region rather than only on clear-cut cases. These cases also stress-test the system's sensitivity near decision boundaries, where a useful assessment tool should still behave meaningfully rather than collapse to one extreme.
 
\section{Experiments}
\label{sec:testing}

This section describes the experimental protocol used to evaluate Hypothetica: the scenarios each idea is run through, the fixed pipeline configuration under which every run is performed, and the metrics used to evaluate the resulting labels. The protocol is designed so that all reported numbers are directly comparable, and so that any observed differences in performance can be attributed to the factor under study -- the corpus, the input length, or an ablated component -- rather than to incidental changes in configuration. To this end, every benchmark case is processed by the same deployed pipeline with identical parameters, and only the single variable of interest is changed within each scenario. The remainder of the section first defines the test scenarios, then lists the default pipeline parameters and the threshold-calibration procedure, and finally specifies the evaluation metrics applied to the resulting predictions.

\subsection{Test Scenarios}

Every benchmark case is evaluated against a single evidence source and passed through the same retrieval, reranking, parsing, and scoring stages used by the deployed application, then mapped to one of the three originality labels. Three scenarios are considered. The \emph{main classification} scenario runs all 135 cases through the full pipeline and compares each predicted label against its gold label, per corpus and per class. The \emph{length-sensitivity} scenario regroups the same predictions by input length tier to isolate the effect of description detail. The \emph{ablation} scenario, run on the OpenAlex corpus, removes or simplifies one component at a time -- query-variant generation, cross-encoder reranking, sentence-level filtering, criterion weighting, and idea enrichment -- to measure each component's contribution. The follow-up question and enrichment step of the deployed system is bypassed in all scenarios, so the reported numbers reflect raw, unenriched idea text.

\subsection{Default Pipeline Parameters}

Table~\ref{tab:params} lists the fixed parameter values used in all experiments. Retrieval and selection parameters control how many candidates reach the scoring phases; scoring parameters determine how per-candidate similarity scores are combined into the final originality verdict.
 
\begin{table}[width=\textwidth,cols=3,pos=ht]
    \caption{Default pipeline parameters used during benchmark evaluation.}
    \label{tab:params}
    \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{3.4cm}p{4.5cm}p{5.8cm}@{}}
        \toprule
        Parameter & Value & Role \\
        \midrule
        Query variants & 4 & Diversified retrieval formulations per idea \\
        Papers per query variant & 150 & Initial retrieval depth per variant \\
        Retained hits per variant & 40 & Candidate retention before deduplication \\
        Embedding top-$k$ & 100 & First-stage semantic retrieval cutoff \\
        Rerank top-$k$ & 20 & Cross-encoder reranking cutoff \\
        Final analyzed candidates & 5 & Documents forwarded to Layer~1 scoring \\
        Likert mapping & $\{1,2,3,4,5\}\rightarrow\{0,0.25,0.5,0.75,1.0\}$ & Criterion score normalization \\
        Criterion weights $(w_p,w_m,w_d,w_c)$ & $(0.15,0.30,0.10,0.45)$ & Problem, method, domain, and contribution weighting \\
        Global aggregation $(w_{\max},w_{\text{mean}})$ & $(0.70,0.30)$ & Max-mean similarity aggregation across candidates \\
        Originality curve exponent $p$ & 1.5 & Nonlinear conversion from similarity to originality \\
        Sentence overlap thresholds & 0.70 / 0.40 & High- and medium-overlap annotation cutoffs \\
        Default score thresholds & $<$40 / 40--69 / $\geq$70 & Uncalibrated label boundaries prior to per-corpus calibration \\
        \bottomrule
    \end{tabular*}
\end{table}

\textbf{Threshold calibration.} Because the raw similarity scale differs across retrieval sources (Section~\ref{sec:crosscorpus}), the results in Section~\ref{sec:results} are reported under per-corpus calibrated thresholds, obtained by selecting the two class boundaries that maximize agreement with the gold labels on each corpus. Only the two thresholds are calibrated; the retrieval and scoring stages are identical across corpora. These figures should therefore be read as achievable accuracy once a deployment is tuned to a target corpus rather than as held-out generalization, and the ceiling this calibration reaches is analyzed in Section~\ref{sec:ceiling}.

\subsection{Evaluation Metrics}

Predicted labels are compared against gold labels using several complementary metrics. \emph{Overall} and \emph{per-class accuracy} measure the fraction of correctly labeled cases in aggregate and within each originality class. Because the classes are balanced, we also report \emph{macro-averaged precision, recall, and F1}, which weight each class equally and expose errors that overall accuracy alone can hide. \emph{Confusion matrices} characterize the direction of errors -- in particular whether misclassifications remain between adjacent classes or jump across the full scale. Accuracy is additionally broken down by length tier, and the ablation scenario is reported as overall OpenAlex accuracy under each configuration. As reference points on this balanced benchmark, a random classifier reaches $33.3\%$ and a majority-class classifier approximately $34\%$.


