\PassOptionsToPackage{hyperfootnotes=false}{hyperref}
\documentclass[a4paper,fleqn]{cas-sc}

\usepackage[authoryear,longnamesfirst]{natbib}

\graphicspath{{figures/}}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

\emergencystretch=3em

% Short title
\shorttitle{Hypothetica}

% Short author
\shortauthors{M. Karakaya et~al.}

% Main title
\title[mode=title]{Hypothetica: A Multi-Agent Framework for
  Literature-Based Novelty Assessment of Research Ideas}
\tnotemark[1]

% Title footnote
\tnotetext[1]{This study was carried out as part of the SENG~491 --
CMPE~491 -- SENG~492 -- CMPE~492 Senior Project I -- Senior Project II
courses, under the supervision of Dr.\ Murat Karakaya at TED University.}

% First author
\author[1]{Harun H\"{u}dai Tan}
\ead{harunhudai.tan@tedu.edu.tr}
\credit{Conceptualization, Methodology, Software, Writing -- Original Draft}

\affiliation[1]{organization={Department of Software Engineering, Ted University},
               addressline={Ziya G\"{o}kalp Caddesi No: 48, Kolej, \c{C}ankaya},
               city={Ankara},
               postcode={06420},
               country={T\"{u}rkiye}}

% Second author
\author[1]{Baran Erol}
\ead{baran.erol@tedu.edu.tr}
\credit{Software, Validation}

% Third author
\author[1]{Ahmet Alp Malko\c{c}}
\ead{ahmetalp.malkoc@tedu.edu.tr}
\credit{Software, Validation}

% Fourth author
\author[2]{Kutay Becerir}
\ead{kutay.becerir@tedu.edu.tr}
\credit{Software, Validation}

\affiliation[2]{organization={Department of Computer Engineering, Ted University},
               addressline={Ziya G\"{o}kalp Caddesi No: 48, Kolej, \c{C}ankaya},
               city={Ankara},
               postcode={06420},
               country={T\"{u}rkiye}}

% Fifth author
\author[1]{Murat Karakaya}
\cormark[1]
\ead{murat.karakaya@tedu.edu.tr}
\credit{Supervision, Writing -- Review \& Editing}

% Corresponding author text
\cortext[1]{Corresponding author: murat.karakaya@tedu.edu.tr}

\begin{abstract}
Researchers often need to spend a lot of time manually reviewing
prior studies to confirm that a proposed project does not repeat
existing work, and this process can be slow.
This paper introduces Hypothetica, an AI-powered system that evaluates
the originality of proposed research ideas by comparing them with
prior art from academic literature, patents, and public code
repositories. The system uses a multi-agent architecture based on large
language models, combined with retrieval-augmented generation (RAG), to
support explainable assessments of originality. The system retrieves
candidate evidence from OpenAlex, arXiv, Google Patents, and GitHub,
then reviews the material across four dimensions: problem similarity,
method similarity, domain similarity, and contribution similarity.
Hypothetica produces a numerical originality score and detailed
reports with evidence-based feedback, mapping sentences in the user's
research idea to related passages in the retrieved sources. The
system's performance was assessed on a 135-case benchmark spanning
three evidence corpora: OpenAlex, Google Patents, and GitHub. In this
evaluation, Hypothetica achieved 75.6\% mean classification accuracy,
with all misclassifications occurring only between adjacent originality
classes. These results indicate that the approach provides a coherent
originality signal across different evidence sources and can help
researchers refine their ideas using clear, practical feedback before
they begin their research.
\end{abstract}

\begin{highlights}
\item Hypothetica assesses research idea originality using a multi-agent RAG pipeline.
\item The system compares ideas against academic literature, patents, and public code repositories.
\item A 135-case benchmark shows 75.6\% mean accuracy with only adjacent-class errors.
\end{highlights}

\begin{keywords}
large language models \sep multi-agent systems \sep research originality \sep retrieval-augmented generation
\end{keywords}

\maketitle

\input{sections/introduction_and_related_work/introduction}
\input{sections/introduction_and_related_work/related_work}
\input{sections/methodology/methodology}
\input{sections/implementation/implementation}
\input{sections/testing_and_results/testing}
\input{sections/testing_and_results/results}
\input{sections/discussion_and_conclusion/discussion}
\input{sections/discussion_and_conclusion/conclusion}

\printcredits

\section*{Declaration of competing interest}

The authors declare that they have no known competing financial interests
or personal relationships that could have appeared to influence the work
reported in this paper.

\section*{Data availability}

The source code and benchmark datasets supporting this study will be made
publicly available in the project repository upon publication.

\bibliographystyle{cas-model2-names}
\bibliography{refs}

\end{document}
